Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-27 demystify.tex
updates

A LoR writer should be someone who knows you well, so that they can provide a strong-supporting and personalized letter for you.  
LoR writers should also have the credibility to convince adcom members. Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
LoR writers are often your research advisors, professors, or supervisors who have worked with you on research projects.

@@

Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs. For example, in PhD programs, you should focus on research areas and advisers (\autoref{sec:schoolsandprofs}), while in MS programs, you might want to consider factors such as location, industry connections, and job placement.

@@

\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather), you should also consider the job opportunities in the area.  If you want to work in the tech industry, you might want to consider MS programs in tech hubs like the Bay Area, Seattle, or Boston. For government jobs, you might want to consider schools in the DC DMV area.  Many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.

\item \textbf{Industry Connections:} Universities often have strong connections with local companies and can help you get internships and jobs. For example, schools in Washington area often have their students working with Microsoft and Amazon while those in the Bay Area have students working with Google and Facebook.  Mason naturally has strong connections with the government and defense contractors and Amazon, whose 2nd HQ is in Arlington.

\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important as you do not need to spend 5--7 years doing research.  Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.  For example, if you're interested in software engineering, you might want to consider schools with strong SWE programs.

\item \textbf{Living Cost:} As MS students are typically not funded, living cost can be important. You should consider the tuition and living expenses of the area. However, I find that in the beginning living costs, e.g., renting, seems scary but students often find way to make it work, e.g., by sharing an apartment with other students.

@@

As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application. It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).

@@

\begin{commentbox}[Vu:]
    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.  Occasionally I was introduced to students by other students or faculty.

    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.  I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.

>>>>

A LoR writer should be someone who knows you well, so that they can provide a strong-supporting and personalized letter for you.  
LoR writers should also have the credibility to convince adcom members. Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
LoR writers are often your research advisers, professors, or supervisors who have worked with you on research projects.

@@

Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs. For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).

@@

\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather, historical city, etc.), you should also consider the job opportunities in the area.  For tech industry, consider MS programs in tech hubs like the Bay Area and Seattle. For government jobs, look at schools in the DC area.  For example, many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.

\item \textbf{Industry Connections:} Universities often have strong connections with local companies and can help you get internships and jobs. For example, Microsoft and Amazon look for students in the Washington area while  Google and Facebook recruit those in the Bay Area.  Mason students natural get jobs from the government, defense contractors, and Amazon, whose 2nd HQ is in Arlington.

\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important.  Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.  For example, if you're interested in software engineering, you might want to consider schools with strong SWE programs.

\item \textbf{Living Cost:} MS students are typically not funded, thus living cost can be a big hassle. You should consider the tuition and living expenses of the area. Note that sometimes living costs, e.g., renting, seems scary at first but students often find way to make it work, e.g., by sharing an apartment with other students.

@@

As discussed in \autoref{sec:research-experience}, having a successful research experience can greatly strengthen a PhD application. Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).

@@

\begin{commentbox}[Vu:]
    I enjoy working with undergrads (e.g., I have mentored over 15 undergrads) and always actively recruit.
    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.  Occasionally I was introduced to students by other students or faculty.\\

    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.  I also have written strong LoRs for them and have helped them get into good PhD programs.
